Bayesian Sparsification of $\cplx$-valued networks

### 2020-01-27 paper/paper.tex
oxford comma, minor rewords

The local reparameterization trick provides both faster computations, and better statistical
efficiency of the SGVB gradient estimator, \citep{wang_fast_2013,kingma_variational_2015}.
We assume that weights $W \in \cplx^{n\times m}$ of a dense linear layer are drawn from a
mean field $\cplx$-Gaussian approximate posterior
\begin{equation}  \label{eq:c-gauss-vi-general}
  W_{ij}
    \sim \mathcal{N}^{\cplx} \bigl(
      \mu_{ij}, \Sigma_{ij}, C_{ij}
    \bigr)
  \,,
\end{equation}
where mean, \textit{variance} and \textit{relation} are given by matrices $
  \mu, \Sigma, C \in \cplx^{n\times m}
$ with $\Sigma_{ij} \in \real$, and $
  \lvert C_{ij} \rvert^2 \leq \Sigma_{ij}
$. Property \eqref{eq:cn-affine} and independence of $W$ imply that for any $x \in \cplx^m$
and $b \in \cplx^n$ the components of $y = W x + b$ are independent with
\begin{equation}  \label{eq:cplx-gauss-trick}
  y_i
    \sim \mathcal{N}^{\cplx}
      \Bigl(
        % e_i^\top \mu x + b_i,
        b_i + \sum_{j=1}^m \mu_{ij} x_j,
        \, \sum_{j=1}^m \Sigma_{ij} \lvert x_j \rvert^2,
        \, \sum_{j=1}^m C_{ij} x_j^2
      \Bigr)
    \,.
\end{equation}
The trick requires three matrix operations: $\mu x + b$, $\Sigma \lvert x \rvert^2$ and
$C x^2$ with the modulus and power applied elementwise.

The trick and dropout can be applied to any layer, the output of which depends linearly
on its parameters, e.g. convolutional layers, affine and bilinear transformations $
  (x, z) \mapsto x^\top W^{(j)} z + b_j
$. Similar to $\real$ case, variational dropout for $\cplx$ convolutions draws independent
kernel $W$ for each spatial patch in the input. The rationale is simpler computations and
variance reduction. Reduced variance stems from decorrelated gradient estimates from
overlapping patches, \citep{kingma_variational_2015}, which allows \eqref{eq:cplx-gauss-trick}
to use $\cplx$ convolutions of elementwise complex squares $x^2$ and amplitudes $\lvert x \rvert^2$
with the relation and variance kernels.
% Weight sharing is essentially pushed to shared variational parameters $\theta$ in the approximation $q_\theta$.
% A convolution is a matrix product of $W$, unrolled into a toeplitz matrix, and im-to-col
% flattened intput $x$

@@

We consider two priors for Bayesian dropout: log-uniform and $\cplx$-Gaussian empirical
Bayes, \citep{kingma_variational_2015,molchanov_variational_2017,kharitonov_variational_2018}.
For a fully factorized approximation $q(\omega)$ and factorized prior belief $\pi(\omega)$,
the divergence term in ELBO \eqref{eq:elbo} is
\begin{equation}  \label{eq:elbo-general-kl-div}
  KL(q \| \pi)
    % = \mathbb{E}_{\oemga \sim q(\oemga)} \log \tfrac{q(\oemga)}{\pi(\oemga)}
    % = \sum_{ij} \mathbb{E}_{\omega_{ij} \sim q(\omega_{ij})}
    %   \log \tfrac{q(\omega_{ij})}{\pi(\omega_{ij})}
    % = \sum_{ij} KL\bigl( q(\omega_{ij}) \| \pi(\omega_{ij}) \bigr)
    = \sum_{ij} \bigl(
        - \mathbb{H}(q(\omega_{ij}))
        - \mathbb{E}_{q(\omega_{ij})} \log{\pi(\omega_{ij})}
      \bigr)
    \,.
\end{equation}
We use approximation \eqref{eq:c-gauss-vi} and omit subscripts ${ij}$ below for brevity.

@@

In this section we will compare the complex variational dropout techniques, discussed
above, on the tasks and datasets, that were studied in prior research on complex-%
valued networks.

@@

\citet{trabelsi_deep_2017} study the applicability of complex-valued networks to image
classification (CIFAR-10, CIFAR-100, reduced train of SVHN), music-transcription
(MusicNet), and Speech Spectrum prediction (TIMIT).
\begin{itemize}
  \item 
\end{itemize}

test on MusicNet, (music transcription [14])

complex high-dimensional time series In-phase Quadrature (IQ)

>>>>

The local reparameterization trick provides both faster computations, and better statistical
efficiency of the SGVB gradient estimator, \citep{wang_fast_2013,kingma_variational_2015}.
Suppose that weights $W \in \cplx^{n\times m}$ of a dense linear layer are drawn from a
mean field $\cplx$-Gaussian approximate posterior
\begin{equation}  \label{eq:c-gauss-vi-general}
  W_{ij}
    \sim \mathcal{N}^{\cplx} \bigl(
      \mu_{ij}, \Sigma_{ij}, C_{ij}
    \bigr)
  \,,
\end{equation}
where mean, \textit{variance} and \textit{relation} are given by matrices $
  \mu, \Sigma, C \in \cplx^{n\times m}
$ with $\Sigma_{ij} \in \real$, and $
  \lvert C_{ij} \rvert^2 \leq \Sigma_{ij}
$. The property \eqref{eq:cn-affine} and independence of elements of $W$ imply that for any
$x \in \cplx^m$ and $b \in \cplx^n$ the components of $y = W x + b$ are independent with
\begin{equation}  \label{eq:cplx-gauss-trick}
  y_i
    \sim \mathcal{N}^{\cplx}
      \Bigl(
        % e_i^\top \mu x + b_i,
        b_i + \sum_{j=1}^m \mu_{ij} x_j,
        \, \sum_{j=1}^m \Sigma_{ij} \lvert x_j \rvert^2,
        \, \sum_{j=1}^m C_{ij} x_j^2
      \Bigr)
    \,.
\end{equation}
The trick requires three matrix operations: $\mu x + b$, $\Sigma \lvert x \rvert^2$ and
$C x^2$ with the modulus and power applied elementwise.

The trick and dropout can be applied to any layer, the output of which depends linearly
on its parameters, e.g. convolutional layers, affine, and bilinear transformations $
  (x, z) \mapsto x^\top W^{(j)} z + b_j
$. Similar to $\real$ case, variational dropout for $\cplx$ convolutions draws independent
realizations of $W$ for each spatial patch in the input. The rationale is simpler computations
and variance reduction. The latter effect stems from decorrelated gradient estimates from
overlapping patches, \citep{kingma_variational_2015}, which allows \eqref{eq:cplx-gauss-trick}
to use $\cplx$ convolutions of elementwise complex squares $x^2$ and amplitudes $\lvert x \rvert^2$
with the relation and variance kernels.
% Weight sharing is essentially pushed to shared variational parameters $\theta$ in the approximation $q_\theta$.
% A convolution is a matrix product of $W$, unrolled into a toeplitz matrix, and im-to-col
% flattened intput $x$

@@

We consider two priors for Bayesian dropout: \verify{log-uniform} and $\cplx$-Gaussian empirical
Bayes, \citep{kingma_variational_2015,molchanov_variational_2017,kharitonov_variational_2018}.
For a fully factorized approximation $q(\omega)$ and factorized prior belief $\pi(\omega)$,
the divergence term in ELBO \eqref{eq:elbo} is
\begin{equation}  \label{eq:elbo-general-kl-div}
  KL(q \| \pi)
    % = \mathbb{E}_{\oemga \sim q(\oemga)} \log \tfrac{q(\oemga)}{\pi(\oemga)}
    % = \sum_{ij} \mathbb{E}_{\omega_{ij} \sim q(\omega_{ij})}
    %   \log \tfrac{q(\omega_{ij})}{\pi(\omega_{ij})}
    % = \sum_{ij} KL\bigl( q(\omega_{ij}) \| \pi(\omega_{ij}) \bigr)
    = - \sum_{ij}
        \mathbb{H}(q(\omega_{ij}))
        + \mathbb{E}_{q(\omega_{ij})} \log{\pi(\omega_{ij})}
    \,.
\end{equation}
We use approximation \eqref{eq:c-gauss-vi} and omit subscripts ${ij}$ below for brevity.

@@

In this section we compare the complex variational dropout techniques discussed above
on the tasks and datasets, that were studied in prior research on complex-valued networks.

@@

\citet{trabelsi_deep_2017} study the applicability of $\cplx$-valued networks to image
classification (CIFAR-10, CIFAR-100, reduced train of SVHN), music-transcription
(MusicNet), and Speech Spectrum prediction (TIMIT).

% complex high-dimensional time series In-phase Quadrature (IQ)
